---
title: "Key taxa summary"
format: html
editor: source
execute:
  echo: false 
  warning: false
  message: false
---


```{r load_packages}
library(tidyverse)
library(ggvenn)
```


```{r load_data}
sb_blast_dat <- paste("/Users/shaman.narayanasamy/Work/Data/key_taxa_ww/extracted sequences of core bacteria/assembly_information/", "sb", "_blast2assembly.tsv", sep = "") %>%  
  read_tsv() %>% 
  filter(`% identity` >= 95)  

swbd_blast_dat <- paste("/Users/shaman.narayanasamy/Work/Data/key_taxa_ww/extracted sequences of core bacteria/assembly_information/", "swbd", "_blast2assembly.tsv", sep = "") %>% 
  read_tsv() %>% 
  filter(`% identity` >= 95)
```

## Key taxa in sludge bulking vs salt water desalination

| System            | BLASTN hits                       | No. of 16S query sequences                              | No. of genome assemblies |
|-------------------|-----------------------------------|-------------------------------------------------------|--------------------------|
| sludge bulking    | `r nrow(sb_blast_dat)`            | `r sb_blast_dat %>% select('query acc.ver') %>% distinct() %>% nrow()` | `r sb_blast_dat %>% select(assembly_id) %>% distinct() %>% nrow()`  |
| swbd              | `r nrow(swbd_blast_dat)`          | `r swbd_blast_dat %>% select('query acc.ver') %>% distinct() %>% nrow()`| `r swbd_blast_dat %>% select(assembly_id) %>% distinct() %>% nrow()` |
| **Total**         | `r nrow(sb_blast_dat) + nrow(swbd_blast_dat)` | `r sb_blast_dat %>% select('query acc.ver') %>% distinct() %>% nrow() + swbd_blast_dat %>% select('query acc.ver') %>% distinct() %>% nrow()` | `r sb_blast_dat %>% select(assembly_id) %>% distinct() %>% nrow() + swbd_blast_dat %>% select(assembly_id) %>% distinct() %>% nrow()` |


Table: Summary of BLASTN hits, distinct (partial) 16S query sequences, and distinct genome assemblies. BLASTN results were filtered based on 95% identity.

```{r assess_overlap}
# Create a named list for the Venn diagram
venn_list <- list(
  sb = sb_blast_dat %>% pull(assembly_id) %>% unique(),
  swbd = swbd_blast_dat %>% pull(assembly_id) %>% unique()
)

# Find the intersection manually to ensure there is overlap
common_ids <- intersect(venn_list$sb, venn_list$swbd)
length(common_ids)  # 0 if there is no overlap

# Generate the Venn diagram with the cleaned data
ggvenn(venn_list, stroke_color = NA)
```

We assess the overlap between the key taxa genomes in the sludge bulking and
salt water desalination biofilm. We find that the there is no overlap between
the key taxa genomes between these systems.


```{r}
# Since thre are no overlaps between the two systems, we can bind the tables together.
blast_dat <- bind_rows(sb_blast_dat, swbd_blast_dat)

blast_dat %>% filter(!is.na(assembly_id)) %>% 
  filter(`% identity` >= 95) %>% 
  select(`query acc.ver`, assembly_id, system, taxonomic_level, taxa) %>% distinct() %>% 
  group_by(`query acc.ver`, taxonomic_level, taxa, system) %>% summarise(no_of_assemblies = n()) %>% 
  arrange(system) %>% 
  knitr::kable(caption = "Summary of no. of genomes representing different taxa and systems")
```


```{r}
genome2cas_system <-
  readLines("/Users/shaman.narayanasamy/Work/Data/key_taxa_ww/summary_data/crisprcasfinder_host_results-v0.2.tsv") %>%  
  tibble(line = .) %>%  
  separate(line, into = paste0("col", 1:20), sep = "\t", fill = "right") %>%  
  select(assembly_id = col1, where(~ !all(is.na(.) | . == ""))) %>% 
  pivot_longer(
    cols = -assembly_id, # Specify columns to pivot into longer format, excluding assembly_id
    names_to = "tmp", # This will be the new column with the names of the pivoted columns
    values_to = "cas_type", # This will be the new column with the values from the pivoted columns
    values_drop_na = TRUE # Drop rows with N/A values in the 'cas_type' column
  ) %>% select(-tmp) %>% 
  separate(assembly_id, into = c("gca", "code")) %>% 
  mutate(assembly_id = paste(gca, code, sep = "_")) %>% 
  mutate(assembly_id = paste(assembly_id, ".1", sep = ""))
```

```{r}
genome2cas_system %>% 
  group_by(assembly_id) %>% 
  summarise(no_of_cas_systems = n()) %>% 
  inner_join(., blast_dat %>% select(assembly_id, taxa, system) %>% distinct, by = "assembly_id") %>% 
  arrange(desc(no_of_cas_systems)) %>% 
  knitr::kable(caption = "Summary of total no. of cas systems per genome")
```
```{r}
genome2cas_system %>% 
  group_by(assembly_id) %>% 
  summarise(no_of_cas_systems = n()) %>% 
  ungroup() %>% 
  inner_join(., blast_dat %>% select(assembly_id, taxa, system) %>% distinct, by = "assembly_id") %>% 
  group_by(taxa) %>% 
  summarise(no_of_cas_systems = sum(no_of_cas_systems)) %>% 
  knitr::kable(caption = "Summary of total no. of cas systems based on taxa")
```

```{r}
genome2cas_system %>% 
  inner_join(., blast_dat %>% select(assembly_id, taxa, system) %>% distinct, by = "assembly_id") %>% 
  group_by(cas_type) %>% 
  summarise(no_of_genomes = n()) %>% 
  arrange(desc(no_of_genomes)) %>% 
  knitr::kable(caption = "Summary of total no. of cas system types accross all genomes")
```
